Multiparty Session Type-safe Web Development in TypeScript

### 2020-01-21 PLACES_2020/paper.tex
PLACES 2020: update author name ordering

\usepackage{listings}
\lstdefinelanguage{Scribble}{
  keywords={module, type, from, as, global,protocol, role, to, choice, at, or, do},
  keywordstyle=\color{blue}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darkgray}\ttfamily,
  stringstyle=\color{purple}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}
\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, class, export, boolean, throw, implements, import, this, const, let},
  keywordstyle=\color{blue}\bfseries,
%  ndkeywords={class, export, boolean, throw, implements, import, this},
%  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darkgray}\ttfamily,
  stringstyle=\color{purple}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}
\lstset{
	basicstyle=\footnotesize\ttfamily,
	numbers=left,
	frame=single,
}
\usepackage{caption}

@@

\title{Multiparty Session Type-safe Web Development in TypeScript}
\author{Anson Miu
\institute{Imperial College London}
\and
Fangyi Zhou
\institute{Imperial College London}
\and
Francisco Ferreira
\institute{Imperial College London}
\and
Nobuko Yoshida
\institute{Imperial College London}
}
\def\titlerunning{MPST-safe Web Development in TypeScript}
\def\authorrunning{A. Miu, F. Zhou, F. Ferreira \& N. Yoshida}
\begin{document}
\maketitle

@@

\paragraph{Components} \dots
% Component = basic building blocks 
% Life cycle
% JSX

\paragraph{Virtual DOM} \dots
% virtual-dom

\paragraph{Uni-directional data flow} \dots
% Parent provides props to child (uni-directional workflow)

\subsubsection{Model types in React}
Although React was originally intended for the \textit{Model-View-Controller} architecture, we use the framework to express the model types approach.

@@

\subsubsection{Affine channel usage}

\paragraph{Repeat use}

@@

\section{Case Study}
\label{section:example}
We apply our framework to implement a web-based implementation of the \textit{Noughts and Crosses} running example in TypeScript; the interested reader can find the full implementation in \cite{NoughtsAndCrosses}. In addition to showing the multiparty session type safety achieved by our generated APIs, we also show that our library design welcomes idiomatic JavaScript practices in the user implementation and is interoperable with common front- and back-end frameworks.

@@

We set up the WebSocket server on top of an Express.js \cite{ExpressJS}application running on Node.js \cite{NodeJS}. We define our own game logic in a \texttt{Board} class to keep track of the game state and expose methods to query the result. This custom logic is integrated into our 
\texttt{Receive\_P1} and \texttt{Receive\_P2} handlers (listing are defined using 
encapsulate our custom game logic in a class

\begin{figure}[!h]
\begin{lstlisting}[language=JavaScript, tabsize=4, title=\texttt{./app.ts}]
const handleP1Move: Receive_P1 = (move: Point) => {
	// User-implemented logic
	board.P1(move)
	if (board.won()) { return [Labels.Lose, move, move] }
	else if (board.draw()) { return [Labels.Draw, move, move] }
	else { return [Labels.Draw, move, handleP2Move] } 
}

// Instantiate session
const webSocketServer: WebSocket.Server = ...
new GameServer(webSocketServer, handleP1Move, handleP2Move)
\end{lstlisting}  
\captionof{lstlisting}{Example fragment of \textit{Noughts and Crosses} \texttt{Svr} implementation.}
\label{lst:svrprotocol}
\end{figure}

@@

\begin{figure}[!h]
\begin{lstlisting}[language=JavaScript, tabsize=4]
export class MakeMove extends S31 {
	render() {
		...
		{board.map((row, x) => (
			board.map((col, y) => {
				const SelectPoint = this.props.Pos('click', event => {
					event.preventDefault()
					return { x, y }
				}
				return <SelectPoint><td>.</td></SelectPoint>
			}
		)}
		...
	}
}
\end{lstlisting}  
\captionof{lstlisting}{Example fragment of \textit{Noughts and Crosses} \texttt{P1} implementation.}
\label{lst:svrprotocol}
\end{figure}

@@

\section{Conclusion and Future Works}

>>>>

\usepackage{listings}
\lstdefinelanguage{Scribble}{
  keywords={module, type, from, as, global,protocol, role, to, choice, at, or, do},
  keywordstyle=\color{blue}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darkgray}\ttfamily,
  stringstyle=\color{purple}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}
\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, class, export, boolean, throw, implements, import, this, const, let, extends},
  keywordstyle=\color{blue}\bfseries,
%  ndkeywords={class, export, boolean, throw, implements, import, this},
%  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darkgray}\ttfamily,
  stringstyle=\color{purple}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}
\lstset{
	basicstyle=\footnotesize\ttfamily,
	numbers=left,
	frame=single,
}
\usepackage{caption}

@@

\title{Multiparty Session Type-safe Web Development in TypeScript}
\author{Anson Miu
\institute{Imperial College London}
\and
Francisco Ferreira
\institute{Imperial College London}
\and
Nobuko Yoshida
\institute{Imperial College London}
\and
Fangyi Zhou
\institute{Imperial College London}
}
\def\titlerunning{MPST-safe Web Development in TypeScript}
\def\authorrunning{A. Miu, F. Ferreira, N. Yoshida \& F. Zhou}
\begin{document}
\maketitle

@@

\paragraph{Components} A component is a reusable UI element which contains its own markup and logic. Components implement a \texttt{render()} function which returns a React element, the smallest building blocks of a React application. Components can keep \textit{state} and the \texttt{render()} function is invoked upon a change of state. A simple counter can be implemented as a component, with the count stored as state, a button which increments the count when clicked and a \texttt{div} that renders the current count. Components can also render other components, which gives rise

\paragraph{Virtual DOM} React components are rendered on a logical abstraction of the DOM, which in turn performs a \texttt{diff} on the current browser DOM and patches the delta accordingly. This allows the programmer to directly specify what should be rendered without having to worry about which elements to append or remove from the browser DOM.

\paragraph{Uni-directional data flow} Parent components pass data to child components as \textit{props}, \dots

% Parent provides props to child (uni-directional workflow)

\subsubsection{Model types in React}
We leverage React to express the model types approach in \cite{MVU2019} rather than the \textit{Model-View-Controller} (MVC) architecture it was intended for.

@@

\subsubsection{Affine channel usage}
A limitation of our browser-side session type encoding is only being able to guarantee that channel resources are used \textit{at most once} as supposed to \textit{exactly once}.

\paragraph{Repeat use} \dots

@@

\section{Case Study}
\label{section:example}

\begin{figure}
\begin{lstlisting}[language=JavaScript, tabsize=4, title=\texttt{./app.ts}]
const handleP1Move: Receive_P1 = (move: Point) => {
	// User-implemented logic
	board.P1(move)
	if (board.won()) { return [Labels.Lose, move, move] }
	else if (board.draw()) { return [Labels.Draw, move, move] }
	else { return [Labels.Draw, move, handleP2Move] } 
}

// Instantiate session
new NoughtsAndCrosses.Svr(webSocketServer, handleP1Move, handleP2Move)
\end{lstlisting}  
\captionof{lstlisting}{Example fragment of \textit{Noughts and Crosses} \texttt{Svr} implementation.}
\label{lst:svrprotocol}
\end{figure}

We apply our framework to implement a web-based implementation of the \textit{Noughts and Crosses} running example in TypeScript; the interested reader can find the full implementation in \cite{NoughtsAndCrosses}. In addition to showing the multiparty session type safety achieved by our generated APIs, we also show that our library design welcomes idiomatic JavaScript practices in the user implementation and is interoperable with common front- and back-end frameworks.

@@

We set up the WebSocket server inside an Express.js \cite{ExpressJS} application on top of a Node.js \cite{NodeJS} runtime. We define our own game logic in a \texttt{Board} class to keep track of the game state and expose methods to query the result. This custom logic is integrated into our 
\texttt{Receive\_P1} and \texttt{Receive\_P2} handlers (listing \ref{lst:svrprotocol}) are defined using

%\dots encapsulate our custom game logic in a class

@@

\begin{figure}[!h]
\begin{lstlisting}[language=JavaScript, tabsize=4]
export class MakeMove extends S31 {
	render() {
		...
		{board.map((row, x) => (
			board.map((col, y) => {
				const SelectPoint = this.props.Pos('click', event => {
					event.preventDefault()
					return { x, y }
				}
				return <SelectPoint><td>.</td></SelectPoint>
			}
		)}
		...
	}
}
\end{lstlisting}  
\captionof{lstlisting}{Example fragment of \textit{Noughts and Crosses} \texttt{P1} implementation.}
\label{lst:clientapp}
\end{figure}

For the sake of sharing the same game client implementation for both players, \cite{NoughtsAndCrosses} uses \textit{higher-order components} (HOC) to build the correct state implementations depending on which player the user chooses to be.

@@

\section{Conclusion and Future Works}
We have presented a MPST-based framework for developing full-stack interactive TypeScript applications with WebSocket communications that conform to a protocol.

Future works include \dots
